---
title: "Reverse Engineering Project"
author: "Caroline and Khushboo"
date: "March 4, 2023"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(dplyr)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

political_donos <- read_csv("data/sports-political-donations.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

donos_clean <- political_donos %>%
  clean_names() %>%
 mutate(amount = as.numeric(gsub("[\\$,]", "",amount)))


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [“Of the 160 owners and commissioners in ESPN’s database, 74 contributed a majority of their publicly searchable funds to Republican campaigns, with 48 donating a majority to Democrats and the rest sending most of their money to causes that aren’t fully aligned with either party.”]

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?] I think the data is rounded a bit based on what we found here. For example, we had 158 donors and he said there were 160. All of our values were slightly less or slightly more. Overall, our reverse engineered sentence is slightly off, and I think this could be because there's a couple of donors missing in our count for some reason? I'm not sure how I'd find who exactly is missing because we got the exact data from GitHub but it would be interesting to go into detail as to where exactly we went off. Perhaps the data has changed since the article was written. After all, it was published in 2020. 

```{r}
# Put code to reverse engineer sentence here

dono_count <- donos_clean %>%
  group_by(owner, party) %>%
  summarise(tot_dono = sum(amount)) %>%
  pivot_wider(names_from = party, values_from = tot_dono) %>%
  replace(is.na(.), 0)

dono_count$top_party = colnames(dono_count)[apply(dono_count,1,which.max)]

count_republican <- dono_count %>%
  filter(str_detect(top_party, "Republican")) %>%
  nrow()

count_democrat <- dono_count %>%
  filter(str_detect(top_party, "Democrat")) %>%
  nrow()

count_other <- dono_count %>%
  filter(!(top_party == "Democrat" | top_party == "Republican")) %>%
  nrow()

donor_count <- donos_clean %>%
  distinct(owner) %>%
  nrow()

count_republican
count_democrat
count_other
donor_count

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: “But while that overall summary is interesting, a more pronounced signal begins to emerge when we look more closely at this group’s larger political donations. There are 40 owners who have donated at least $100,000 to Republican causes over the six-year period, compared to 23 owners topping $100,000 to Democrats, helping lead to these splits:”

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?] We found that the amount of owners that gave money to democrats was 22, and the amount donating to republicans was 37. This is slightly off, considering he found there to be 40 owners for Republican donations and 23 owners for Democrat donations for at least 100,000 dollars. 

```{r}
# Put code to reverse engineer sentence here

larger_donos <- donos_clean %>%
  group_by(owner, party) %>%
  summarise(tot_dono = sum(amount)) %>%
  filter(tot_dono >= 100000)

large_republican <- larger_donos %>%
  filter(party == "Republican") %>%
  nrow()

large_democrat <- larger_donos %>%
  filter(party == "Democrat") %>%
  nrow()

large_republican
large_democrat

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: “Overall, 51.7 percent of all contributions by WNBA owners went to Democrats, with 42.3 percent going to Republicans. And of the $1.3 million contributed by WNBA owners to Republicans, 65.5 percent comes from Kelly Loeffler, the Republican from Georgia who is co-owner of the Atlanta Dream.”

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?] We found that we agreed with the author's findings with accuracy. Our percentages mirrored that of the sentence exactly. 

```{r}
# Put code to reverse engineer sentence here

wnba_donos <- donos_clean %>%
filter(str_detect(league, "WNBA")) %>%
group_by(owner, party) %>%
  summarise(total_donos = sum(amount))

wnba_by_party <- wnba_donos %>%
  group_by(party) %>%
  summarise(donos = sum(total_donos)) %>%
  mutate(percent = donos/(sum(donos))*100)

wnba_republican <- wnba_donos %>%
  filter(party == "Republican") %>%
  summarise(total_donos) %>%
  mutate(percent = total_donos/(sum(total_donos))*100)

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: 

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-