---
title: "lab_03"
author: "Caroline Koutsos"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse

# Load janitor
```

## Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans and, using janitor, change the columns so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows

# Load in the Maryland grants and loan data

```

```{r}
payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2023.csv")
```



```{r}
# cleaning function
cleaned_payments <- payments %>%
  clean_names ()
#display the cleaned data set
cleaned_payments

```


## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed. Write code to find rows where the vendor name contains the county's school system, starting with a broad search and then narrowing it down. When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

A1: about 27 but there definitely could be more that I'm simply not able to find 

```{r}
#cleaning function 
pg_schools <- cleaned_payments %>%
  filter(vendor_name == "PRINCE GEORGES COUNTY PUBLIC SCHOOL" | vendor_name == "PRINCE GEORGES COUNTY PUBLIC SCHOOLS")

```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? What do you notice about the years listed?

A2: Largest total amount is 2021. However, 2009 and 2008 are very close together in total amount. 2010 is very close to those as well. 2021 is an outlier in this case. In 2022, the fiscal budget was 2.7 billion according to WTOP news, so I assume its increasing because of mental health initiatives, COVID, and other things.  

```{r}
pg_schools %>%
  group_by(fiscal_year) %>%
  summarise(total_amount = sum(amount)) %>%
    arrange(desc(total_amount))
```
Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received. Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

A3. Well first of all, some of the zip codes only have 4 digits while some have 5. This might mean we're missing a leading zero, so we can use the string functions by making a new column for our zip codes and use mutate/ string sub to trim the zip code to five characters. We can solve the zero by using case_when and paste0 to any zip that's less than 5. One vendor name issue is A&I INC whereas the other is A & I INC. So we should combine them by having them not have a space. Also A MORTON THOMAS & ASSOC INC is shortened in one, and ASSOCIATES in the other two, so I'd filter to see where they all are and then use get dupes maybe to see the duplicates? I'd definitely use get dupes for Turner Construction Company though because they're the same and have different zip codes and total amounts so I would get rid of duplicate rows by using string and distinct which would get rid of the duplicates. 

```{r}
cleaned_payments %>%
  filter(agency_name == "CAPITAL PROJECTS- UNIVERSITY OF MARYLAND") %>%
  group_by(vendor_name, vendor_zip) %>%
  summarise(total_amount = sum(amount))
```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. Then create a total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

A4: So over time the amount of money towards stem cell research is decreasing by a lot and then slightly going up again. In 2022, it is around 7,000,000 but in 2009 it is around 14,000,000. 

```{r}
grants <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv")
# cleaning function
cleaned_grants <- grants %>%
  clean_names ()
#display the cleaned data set
cleaned_grants %>%
  mutate(description = str_to_upper(description)) %>%
filter(str_detect(description , "STEM CELL")) %>%
group_by(fiscal_year) %>%
summarise(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
```
