---
title: "lab_10"
author: "caroline koutsos"
date: "2023-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* Our usual libraries, plus some of the map-specific ones from the pre_lab. You'll also be fetching Census data in Q3. Read the questions carefully!

put geometry first in join

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Load libraries
library(tidyverse)
library(sf)
library(tigris)
library(janitor)
library(tidycensus)
library(lubridate)
library(ggplot2)
library(ggthemes)

```

Let's explore Maryland elections, power outages and electric vehicle data and make some maps! 

## Questions 

**Q1.** Maryland voters made history in November by electing the state's first two African-American statewide office holders: Wes Moore as governor and Anthony Brown as attorney general. You want to compare the two's electoral performances. Using the `md_statewide_county.csv` file in the data folder, calculate the total number of votes for both the governor's race and the attorney general's race, then percentages for both Moore and Brown _based on the total of their respective contests_. Then calculate the difference between their percentages.

Get the geometry for Maryland's counties using tigris and, joining that dataframe to your results dataframe, make a county-level map of the differences between Moore and Brown's percentages. What is the story here, and what is the county with the most interesting results?

**A1.** In Baltimore City and Prince George's County, Anthony Brown had a higher percentage than Wes Moore of the total votes for their respective contests. Conversely, Worcester, Kent, and Carroll all have the highest difference in voting percentage between Moore and Brown in Moore's favor. Prince George's County is 59% black and 12% white whereas Carroll County is 86% white and 3% black. This could show the significance of representation in Maryland Government for counties that have a larger black population. 

```{r}
md_statewide_county <- read_csv("data/md_statewide_county.csv")

md_statewide_county <- md_statewide_county %>%
  mutate(totalvotesgov = Cox + Moore + Lashar + Wallace + Harding + Sefcik + GovWritein) %>%
  mutate(totalvotesag = Peroutka + Brown + AgWritein) %>%
  mutate(pctmoore = Moore/totalvotesgov*100) %>%
  mutate(pctbrown = Brown/totalvotesag*100) %>%
  mutate(pctmoorebrown = pctmoore - pctbrown)

md_counties <- counties %>%
  filter(STATEFP == "24")

md_statewide_county <- md_statewide_county %>% mutate(GEOID = as.character(GEOID))

moore_brown_plus_counties <- md_counties %>% left_join(md_statewide_county, by=c("GEOID"))

ggplot() +
  geom_sf(data=moore_brown_plus_counties, aes(fill=pctmoorebrown)) +
  scale_colour_viridis_b(option="magma") +
  theme_minimal()
```

**Q2.** Make a county-level map to explore the number of customers impacted by power outages in Maryland. Load the `county_outages_march2023.csv` file in the data folder and calculate a rate of customers with outages per 10,000 total customers for each county. Then join that dataframe to your Maryland counties dataframe you made in Q1 and create a county-level map of the results; you may need to tweak the fill scale to make the distinctions clearer. What's the potential story here? Given access to other months' outage data, what would your next steps be in reporting a story?

**A2.** Talbot County has a higher rate of power outages per capita than any other county by far during the month of March. I feel like the story here has to do with Talbot County infrastructure and how they might be unable to keep the power during the storm season of March? I would want to see the summer months and how summer storms affect Talbot county. 

```{r}
outages_by_county <- read_csv("data/county_outages_march2023.csv")

outages_by_county <- outages_by_county %>%
  mutate(per_capita = outage_customers/total_customers*10000)

county_outages_md <- md_counties %>% left_join(outages_by_county, by=c( "NAME" = "county"))

ggplot() +
  geom_sf(data=county_outages_md, aes(fill=per_capita)) +
  scale_colour_viridis_b(option="magma", trans = "log") +
  theme_minimal()

```

**Q3.** You've been assigned a story about the use of electric vehicles in Maryland and you want to make a map of the prevalence of electric vehicles in each county. Load the `ev_2020_2023.csv` file, which has the number of electric and hybrid vehicles per county from July 2020 and March 2023 and calculate a percentage change for each county (remember, that's `new-old/old`). Using the Maryland counties dataframe you made for Q1, make a county map showing the percentage change.What counties stand out to you? What questions do you have about the changes you're seeing?

Next, use tidycensus to retrieve county population data for Maryland from 2021 (you did this in the pre_lab) and join it to your vehicles dataframe. HINT: you may not need to join it on the county name, and you should already have the geometry!

Then, calculate a per capita rate of electric & hybrid vehicles using the March 2023 figure and the population estimate. You will need to choose the "per capita" value.

Finally, make a second county map showing the per capita rate. What counties stand out on this map? How is it different from the first one you made?

**A3.** 1st Map: Caroline County most 
2nd Map: Howard and Montgomery counties with the most

So for the first map, definitely Caroline County has the most electric vehicles that have changed over time. All of the counties have at least changed by 100% too, so a story could revolve around the initiatives towards sustainability in Maryland this past year. 

When it comes to the per capita rate, Montgomery and Howard Counties stand out as the most per every 10,000 people. I think this could be more indicative of how many people are actually getting electric vehicles, especially since Caroline County had only 42% per capita rate. 

Mainly, this might have to do with the Clean Cars II Rule Wes Moore instituted in 2023. 

Percent change: rate change two points in time 
per capita: the rate per every 10,000 people

```{r}
ev <- read_csv("data/ev_2020_2023.csv")

ev <- ev %>%
  mutate(percent_change = ((march_2023 - july_2020)/(july_2020))*100)

ev_counties <- md_counties %>% left_join(ev, by=c( "NAME" = "county"))

ggplot() +
  geom_sf(data=ev_counties, aes(fill=percent_change)) +
  scale_colour_viridis_b(option="magma", trans = "log") +
  theme_minimal()
  

```

```{r}

md_county_population <- get_acs(geography = "county",
              variables = c(population = "B01001_001"),
              year = 2021,
              state = "MD")

md_pop_with_ev <- ev_counties %>% left_join(md_county_population, by=c("GEOID")) %>%
  mutate(per_capita = march_2023/estimate*10000)

ggplot() +
  geom_sf(data=md_pop_with_ev, aes(fill=per_capita)) +
  scale_colour_viridis_b(option="magma", trans = "log") +
  theme_minimal()

```

